Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2022-01-21 Manuscript/Manuscript.Rmd
check assumptions Heckman model

```{r pop_norm_eval}
# plot(imp_norm)
plot_box(imp_norm_2l, x = "popular", strip = TRUE)

@@

```{r pop_pan_eval}
# plot(imp_pan)
plot_box(imp_pan_2l, x = "popular", strip = TRUE)
ICCs <- ICCs %>% mutate(
           pan = c(icc(popular ~ as.factor(class), complete(imp_pan_2l)), 
                    icc(popteach ~ as.factor(class), complete(imp_pan_2l)), 
                    icc(texp ~ as.factor(class), complete(imp_pan_2l)))
           )
ICCs
```

@@

Here we can set `maxit` to 1...

>>>>

```{r pop_norm_eval, echo=FALSE}
# plot(imp_norm)
plot_box(imp_norm_2l, x = "popular", strip = TRUE)

@@

```{r pop_pan_eval, echo=FALSE}
# plot(imp_pan)
plot_box(imp_pan_2l, x = "popular", strip = TRUE)
ICCs <- ICCs %>% mutate(
           pan = c(icc(popular ~ as.factor(class), complete(imp_pan_2l)), 
                    icc(popteach ~ as.factor(class), complete(imp_pan_2l)), 
                    icc(texp ~ as.factor(class), complete(imp_pan_2l)))
           )
ICCs
```

@@

From the missing data pattern we see that we can set `maxit` to 1, since there is only one variable with missingness.

```{r include=FALSE}
# inclusion restriction variable 
mod_main <- glm(hiv ~ ., family = "binomial", data = hiv) %>% 
  broom::tidy(conf.int = TRUE) %>% 
  .[7, c(2,6,7)] %>% 
  round(., 3)
mod_select <- glm(is.na(hiv) ~ ., family = "binomial", data = hiv) %>% 
  broom::tidy(conf.int = TRUE) %>% 
  .[7, c(2,6,7)] %>% 
  round(., 3)
```

The inclusion restriction variable should be a predictor of the missingness indicator for the variable of interest, but *not* of the actual value of the variable of interest. In this case, the data were simulated to adhere to this requirement. Namely, $\beta_{smoke}$ = `r mod_main$estimate`, 95% CI [`r paste0(mod_main$conf.low, ", ", mod_main$conf.high)`] for the analysis model (`formula = hiv ~ .`), and $\beta_{smoke}$ = `r mod_select$estimate`, 95% CI [`r paste0(mod_select$conf.low, ", ", mod_select$conf.high)`] for the selection model (`formula = is.na(hiv) ~ .`). This means the assumptions for the Heckman model are met.

### 2022-01-21 Manuscript/Manuscript.tex
check assumptions Heckman model

\begin{CodeChunk}
\begin{CodeInput}
R> # plot(imp_norm)
R> plot_box(imp_norm_2l, x = "popular", strip = TRUE)
\end{CodeInput}

@@

\begin{CodeInput}
R> ICCs <- ICCs %>% mutate(
+            norm = c(icc(popular ~ as.factor(class), complete(imp_norm_2l)), 
+                     icc(popteach ~ as.factor(class), complete(imp_norm_2l)), 
+                     icc(texp ~ as.factor(class), complete(imp_norm_2l)))
+            )
R> ICCs
\end{CodeInput}
\begin{CodeOutput}
      vars       CCA   ignored predictor      norm
1  popular 0.3989463 0.4235643 0.3904372 0.4165006
2 popteach 0.3605902 0.3605902 0.3605902 0.3605902
3     texp 1.0000000 1.0000000 1.0000000 1.0000000
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> # plot(imp_pan)
R> plot_box(imp_pan_2l, x = "popular", strip = TRUE)
\end{CodeInput}

@@

\begin{CodeInput}
R> ICCs <- ICCs %>% mutate(
+            pan = c(icc(popular ~ as.factor(class), complete(imp_pan_2l)), 
+                     icc(popteach ~ as.factor(class), complete(imp_pan_2l)), 
+                     icc(texp ~ as.factor(class), complete(imp_pan_2l)))
+            )
R> ICCs
\end{CodeInput}
\begin{CodeOutput}
      vars       CCA   ignored predictor      norm       pan
1  popular 0.3989463 0.4235643 0.3904372 0.4165006 0.3827595
2 popteach 0.3605902 0.3605902 0.3605902 0.3605902 0.3605902
3     texp 1.0000000 1.0000000 1.0000000 1.0000000 1.0000000
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{itemize}
\tightlist
\item
  \texttt{region} Cluster variable,
\item
  \texttt{hiv} HIV diagnosis (0=no, 1=yes),
\item
  \texttt{age} Age of the patient,
\item
  \texttt{marital} Marital status,
\item
  \texttt{condom} Condom use during last intercourse,
\item
  \texttt{smoke} Smoker (levels).
\end{itemize}

@@

Here we can set \texttt{maxit} to 1\ldots{}

>>>>

@@

\begin{CodeOutput}
      vars       CCA   ignored predictor      norm
1  popular 0.3989463 0.4235643 0.3904372 0.4165006
2 popteach 0.3605902 0.3605902 0.3605902 0.3605902
3     texp 1.0000000 1.0000000 1.0000000 1.0000000
\end{CodeOutput}
\end{CodeChunk}

@@

@@

\begin{CodeOutput}
      vars       CCA   ignored predictor      norm       pan
1  popular 0.3989463 0.4235643 0.3904372 0.4165006 0.3827595
2 popteach 0.3605902 0.3605902 0.3605902 0.3605902 0.3605902
3     texp 1.0000000 1.0000000 1.0000000 1.0000000 1.0000000
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{itemize}
\tightlist
\item
  \texttt{region} Cluster variable,
\item
  \texttt{hiv} HIV diagnosis (0=no, 1=yes),
\item
  \texttt{age} Age of the patient,
\item
  \texttt{marital} Marital status,
\item
  \texttt{condom} Condom use during last intercourse,
\item
  \texttt{smoke} Smoker (levels; inclusion restriction variable).
\end{itemize}

There is a specific MNAR mechanism.

\begin{itemize}
\tightlist
\item
  main model = analysis model = predicting the prevalence of hiv
\item
  people who have hiv are more reluctant to taking the test, not to
  reveal their status
\item
  solution: take variables from analysis model like condom use, marital
  status, age, etc. that relate to hiv prevalence
\item
  exclusion restriction variable = instrumental variable (from
  economics) = e.g.~how kind the nurse was = smoke in this dataset =
  should be correlated to the missingness indicator, but \emph{not} the
  variable itself
\item
  include these variables in the selection equation
\item
  take the missingness indicator of the variable of interest for the
  selection equation as DV
\item
  heckman models estimate BOTH selection and mean models simultaneously,
  because linked through the error term
\item
  if the errors are not correlated, the models are not linked at all
  -\textgreater{} not MNAR, but just MAR
\item
  if the rho is significantly different from zero, the errors are
  correlated and the models are related so we have MNAR
\item
  now, we add clustering
\item
  copulas = estimate 2 equations separately and use another equation to
  link them
\item
  for each of the clusters, est the copula separately, so we have all
  coeffs for each cluster, incl the rho and sigma (the errors)
\item
  put together all of the estimates through a meta-analysis model to get
  population distribution of the estimates
\item
  use the population parameter estimates to draw marginal parameters
  from the distribution to get a cluster/study parameter
\item
  with systematical miss we can est the values of the imputed values
  based on the marginal distribution
\item
  optional to add shrinkage: the marginal in account as well
\item
  the hiv data was simulated by the copula function/model developer
\item
  we could add a systematically missing variable in the data, because
  these are all sporadically
\end{itemize}

@@

From the missing data pattern we see that we can set \texttt{maxit} to
1, since there is only one variable with missingness.

The inclusion restriction variable should be a predictor of the
missingness indicator for the variable of interest, but \emph{not} of
the actual value of the variable of interest. In this case, the data
were simulated to adhere to this requirement. Namely, \(\beta_{smoke}\)
= -0.064, 95\% CI {[}-0.256, 0.126{]} for the analysis model
(\texttt{formula\ =\ hiv\ \textasciitilde{}\ .}), and \(\beta_{smoke}\)
= -0.265, 95\% CI {[}-0.422, -0.11{]} for the selection model
(\texttt{formula\ =\ is.na(hiv)\ \textasciitilde{}\ .}). This means the
assumptions for the Heckman model are met.
